---
title: "Data_Science_5_a_1"
output: html_document
---
Set Working directory
```{r}
setwd("c:/R/c5p1/RepData_PeerAssessment1")
```




Unzipping the base data
```{r}
unzip("activity.zip")
```

Read in the Data csv data
```{r}
a<-read.csv("activity.csv")
```

Loading the Correct R Packages for the Data Analysis
```{r}
library('dplyr')
```


This adds the number of steps for each day and then creates a hisogram of the
steps each day. 
```{r}
a1<-aggregate(steps ~date,data=a,sum)
hist(a1$steps)
```


Determine the mean and median of the steps taken in a day.  
Ignoring the NAs.
```{r}
a2<-mean(a1[,2])
a3<-median(a1[,2])
```
The mean is 10766.2 and the median is 10765


Making a plot of the daily activity pattern
```{r}
## Find the average for each interval
a4<-aggregate(steps ~interval,data=a,mean)
plot(a4,type="l",main="Average Number of Steps for each Interval")
## Find the interval with the maximum number of steps
a5<-max(a4[,2])
## find the interval number where the maximum occurs
a6<-a4[match(a5,a4[,2]),1]
```
The maximum occurs at in interval 835.


Determining the number of NAs in the data set
```{r}
a7<-sum(is.na(a[,1]))
```
There are 2304 NAs in the data set. The NAs will be replaced with the 
average value for the time interval from the entire datasets.

```{r}
## number of measurements
b1<-length(a[,1])

## creating the array that will have values replaced in
b=a

## Replacing NAs
for (i in 1:b1) {
        if(is.na(b[i,1])){
                b[i,1]<-a4[match(b[i,3],a4[,1]),2]
                }
           
        }
## finding the total number of steps per day 
b2<-aggregate(steps ~date,data=b,sum)

## Create a hisogram of the number of steps per day
hist(b2$steps)

## Mean
b3<-mean(b2[,2])
## Median
b4<-median(b2[,2])


```
The new median and average are the same at 10766.2.
THe new median and average are for all practical purposes the same. The average
is the same, but the median is a bit diffrent.  After imputation, there are 8
additional days of data, so all the bins of the histogram increase in frequency.
The distribution of the histogram does not appear to change significantly.

Creating the factor variable for weekday or Weekend for the dataset.

```{r}
## week day vector
c1<-weekdays(as.Date(as.character(b[,2])))

### Replacing the character week days with weekend or weekday factors 
i=0
for (i in 1:b1){
        if(c1[i]=="Saturday"|c1[i]=="Sunday"){c1[i]="Weekend"}
        else {c1[i]="Weekday"}
        
        }
c2<-as.factor(c1)
d<-cbind(b,c2)
```


Now sorting the dataset into weekdays and weekends
```{r}
d1<-filter(d,c2=="Weekday")
d2<-filter(d,c2=="Weekend")

```

Finding the average number of steps per interval for weekdays and weekends
```{r}
e1<-aggregate(steps ~interval,data=d1,mean)
e2<-aggregate(steps ~interval,data=d2,mean)
```

Plotting the graph with both weekday and Weekend data
```{r}
par("mfrow"=c(2,1))
plot(e1,type="l",xlab="Interval in Minutes",ylab="Steps")
title("Weekday Data")
plot(e2,type="l",xlab="Interval in Minutes",ylab="Steps")
title ("Weekend Data")

```

