---
title: "Data_Science_5_a_1"
output: html_document
---
Setting Working directory
```{r echo=FALSE}
setwd("c:/R/c5p1/RepData_PeerAssessment1")
```




Unzipping the base data
```{r echo=FALSE}
unzip("activity.zip")
```

Reading in the Data csv data
```{r echo=FALSE}
a<-read.csv("activity.csv")
```

Loading the Correct R Packages for the Data Analysis
```{r echo=FALSE}
library('dplyr')
```


Adding the number of steps for each day 
```{r echo=FALSE}
a1<-aggregate(steps ~date,data=a,sum)
```

Creating a histogram of the steps each day. 
```{r echo=FALSE}
hist(a1$steps, xlab="Steps per Day",ylab="Days",main="Histogram of Steps per Day")
```




Determine the mean and median of the steps taken in a day and ignoring the NAs.
```{r echo=FALSE}
a2<-mean(a1[,2])
a3<-median(a1[,2])
```

```{r echo=FALSE}
print("Mean")
a2
print ("Median")
 a3
```











Making a plot of the daily activity pattern
```{r echo=FALSE}
## Find the average for each interval
a4<-aggregate(steps ~interval,data=a,mean)
plot(a4,type="l",main="Average Number of Steps for each Interval")
## Find the interval with the maximum number of steps
a5<-max(a4[,2])
## find the interval number where the maximum occurs
a6<-a4[match(a5,a4[,2]),1]
```
The maximum Number of Steps occurs at interval
```{r echo=FALSE}
a6
```


Determining the number of NAs in the data set
```{r echo=FALSE}
a7<-sum(is.na(a[,1]))
```
Number of NAs

```{r echo=FALSE}
a7
```

The NAs will be replaced with the average value for the time interval from the entire dataset.

```{r echo=FALSE}
## number of measurements
b1<-length(a[,1])

## creating the array that will have values replaced in
b=a

## Replacing NAs
for (i in 1:b1) {
        if(is.na(b[i,1])){
                b[i,1]<-a4[match(b[i,3],a4[,1]),2]
                }
           
        }
## finding the total number of steps per day 
b2<-aggregate(steps ~date,data=b,sum)

## Create a hisogram of the number of steps per day
hist(b2$steps, xlab="Steps per Day",ylab="Days",Main="Number of Steps per Day")
```









Determining the Mean and Median of the new dataset.
``` {r echo=FALSE}
## Mean
b3<-mean(b2[,2])
## Median
b4<-median(b2[,2])

```
Mean with imputed Values:
```{r}
b3
```
Median with imputed Values:
```{r}
b4
```

The new median and average are for all practical purposes the same. The average
is the same, but the median is a bit diffrent.  After imputation, there are 8
additional days of data, so all the bins of the histogram increase in frequency.
The distribution of the histogram does not appear to change significantly.

Creating the factor variable for weekday or Weekend for the dataset.

```{r echo=FALSE}
## week day vector
c1<-weekdays(as.Date(as.character(b[,2])))

### Replacing the character week days with weekend or weekday factors 
i=0
for (i in 1:b1){
        if(c1[i]=="Saturday"|c1[i]=="Sunday"){c1[i]="Weekend"}
        else {c1[i]="Weekday"}
        
        }
c2<-as.factor(c1)
d<-cbind(b,c2)
```


Now sorting the dataset into weekdays and weekends
```{r echo=FALSE}
d1<-filter(d,c2=="Weekday")
d2<-filter(d,c2=="Weekend")

```

Finding the average number of steps per interval for weekdays and weekends
```{r echo=FALSE}
e1<-aggregate(steps ~interval,data=d1,mean)
e2<-aggregate(steps ~interval,data=d2,mean)
```

Plotting the graph with both weekday and Weekend data
```{r echo=FALSE}
par("mfrow"=c(2,1))
plot(e1,type="l",xlab="Interval in Minutes",ylab="Steps")
title("Weekday Data")
plot(e2,type="l",xlab="Interval in Minutes",ylab="Steps")
title ("Weekend Data")

```

